---
title: Lambs carcass composition
author:
   - Vasco Cadavez
   - Ursula Gonzales Barron
date: 2024/04/11
output: 
  rmarkdown::html_vignette:
     number_sections: yes
     toc: yes
vignette: >
  %\VignetteIndexEntry{Lambs carcass composition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage{caption, amssymb, amsmath}
bibliography: ../inst/REFERENCES.bib
csl: ../inst/apa.csl
---

## Introduction

In this tutorial we will explain how to use the `pasagropasr` R package developed to accommodate all the datasets resulted from the reseach activities of the PAS-AGRO-PAS project. The first step is load the `pasagropasr`, and all the packages that we need for the data analysis.

```{r}
library(pasagropasr)
library(ggplot2)
```

## Loading data

To conduct statistical analyses, we need to import data into R working environment. The `predmicror` package has incorporated datasets, and we can use the `data()` function to load the example datasets to the working environment. Thus, we will start by fitting a full growth model to experimental data, thus we load the `lambs.rda` dataset which is part of the `predmicror` package.

```{r}
library(readxl)
lambBreeds <- readxl::read_xlsx("../inst/extdata/lambBreeds.xlsx", sheet=1)
```

Now the dataset is available in the `R` environment, and we can print the entire `dataset` by typing `lambBreeds` or take a look to the first five lines by using the `head()`.

```{r}
lambBreeds
head(lambBreeds)
```

For data outer to `predmicror`package, usually in .csv format, which are flat text files we use the `read.csv()` function to import the CSV file into R environment. Before load a dataset, its good practice to assure that the dataset is located in the working directory, thus to import a CSV file called `growthfull.csv` into the R environment we do it with the next code section.

```{r, eval=TRUE}
library(readxl)
lambBreeds <- readxl::read_xlsx("../inst/extdata/lambBreeds.xlsx", sheet=1)
```

We have the dataset in the R environment, thus we can start checking the data proprieties. For example, the `str()` function gives information considering the structure of the variables (numeric, integer, etc.), and the `names()` function show us the variables names.

```{r}
str(lambBreeds)
names(lambBreeds)
```

## Plotting data

To check the relationship between `Time` and `lnN`, we might use the function `plot()`, and we can check the data by visual inspection.

```{r}
plot(MUS ~ KKCF,
  data = lambBreeds,
  xlab = "%", ylab = "%"
)
```

