---
title: Lambs growth
author:
   - Vasco Cadavez
   - Ursula Gonzales Barron
date: 2024/04/11
output: 
  rmarkdown::html_vignette:
     number_sections: yes
     toc: yes
vignette: >
  %\VignetteIndexEntry{Lambs growth }
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage{caption, amssymb, amsmath}
bibliography: ../inst/REFERENCES.bib
csl: ../inst/apa.csl
---

## Introduction

In this tutorial we will explain how to fit primary growth models to experimental data. For this lab, we will use the `predmicror` R package developed to accommodate all the functions we are being using in predictive microbiology workshops. The `predmicror` package is outside the base R functions, and the first step is load the `predmicror`, package with the principal predictive microbial growth models, and `gslnls` which is a package to fit non-linear models using non-linear least squares.



```{r}
library(pasagropasr)
library(ggplot2)
```

## Loading data

To conduct statistical analyses, we need to import data into R working environment. The `predmicror` package has incorporated datasets, and we can use the `data()` function to load the example datasets to the working environment. Thus, we will start by fitting a full growth model to experimental data, thus we load the `lambs.rda` dataset which is part of the `predmicror` package.

```{r}
data(lambs)
```

Now the dataset is available in the `R` environment, and we can print the entire `dataset` by typing `growthfull` or take a look to the first five lines by using the `head()`.

```{r}
lambs
head(lambs)
```

For data outer to `predmicror`package, usually in .csv format, which are flat text files we use the `read.csv()` function to import the CSV file into R environment. Before load a dataset, its good practice to assure that the dataset is located in the working directory, thus to import a CSV file called `growthfull.csv` into the R environment we do it with the next code section.

```{r, eval=FALSE}
lambs <- read.csv("../inst/extdata/lambs.csv", sep = ",", header = TRUE)
```

We have the dataset in the R environment, thus we can start checking the data proprieties. For example, the `str()` function gives information considering the structure of the variables (numeric, integer, etc.), and the `names()` function show us the variables names.

```{r}
str(lambs)
names(lambs)
```

## Plotting data

To check the relationship between `Time` and `lnN`, we might use the function `plot()`, and we can check the data by visual inspection.

```{r}
plot(kkcf ~ PCQ,
  data = lambs,
  xlab = "cm", ylab = "cm"
)
```

